Isotropic Stellar Obliquities of Misaligned Planetary Systems

### 2022-12-31 src/tex/ms.tex
[showyourwork] overleaf sync

Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation environment and dynamical evolution of a planetary system. In exoplanet observations, only sky-projected stellar obliquity $\lambda$ can be measured. To find the stellar obliquity from the projected obliquity, information on stellar inclination $i_\star$ is required.
Here we show that, while it is important to know the stellar inclination to tightly constrain the stellar obliquity of an individual system, stellar inclination is not necessarily required for population stellar obliquity inference. The stellar obliquity distribution can be inferred purely from the projected stellar obliquities.
We introduce a flexible hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems from projected stellar obliquities, stellar inclinations, or both.
Applying the framework to all systems with projected stellar obliquity measurements, which are mostly Hot Jupiter systems, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees. Misaligned systems have nearly isotropic stellar obliquity distribution, at odds with previous observations.

@@

Stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$. 
The angle is an important trace of the formation environment and dynamical evolution of a planetary system. The stellar obliquity evolution can be summarized into three phases. First, the formation and evolution of a protoplanetary disk around its host star determine the primordial stellar obliquity. Second, dynamical evolution in the planetary system excites the mutual inclinations between planets or stellar companions and further sculpts the stellar obliquity. Lastly, tidal realignment of the host star's spin axis could reduce the stellar obliquity.

Stellar obliquity is an important tracer to understanding the formation of close-in planetary systems, particularly of Hot Jupiters. Various proposed origin channels of Hot Jupiters make predictions on the stellar obliquity distribution of the Hot Jupiters, assuming that the primordial stellar obliquity is close to zero.

In reality, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured. Stellar obliquity $\psi$ can be computed from $\lambda$ as \citep[e.g.,][]{Fabrycky09}:
\begin{equation}\label{eqn:psi}
    \cos{\psi} = \sin{i_\star}\sin{i_{\rm orb}}\cos{\lambda} + \cos{i_\star}\cos{i_{\rm orb}},
\end{equation}
where $i_{\rm orb}$ is the angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the angle between $\bf{n}_{\star}$ and the observer's line of sight.
Since the sky-projected stellar obliquities can be measured only on transiting exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
To constrain the 3D stellar obliquity $\psi$, the measurements of the sky-projected obliquity $\lambda$ and stellar inclination $i_\star$ are required. The stellar inclination can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology. Since the stellar inclination measurement is degenerate between $i_\star$ and $180\degr - i_\star$, $i_\star$ is usually assumed to be less than $90\degr$.

Hot Jupiter systems' projected stellar obliquities are distributed widely but with a concentration at zero.

For systems without measured $i_\star$, it is still possible to infer the stellar obliquity $\psi$ from the sky-projected obliquity $\lambda$ by assuming the stellar inclination $i_\star$ follows the isotropic distribution.

In this work, we find an approach to infer $\psi$ distribution from $\lambda$ distribution.

In this work, we build a hierarchical Bayesian framework to constrain the stellar obliquity distribution of exoplanetary systems, particularly of misaligned systems.
In Section~\ref{sec:jacobian},
In Section~\ref{sec:hbm}, we describe our hierarchical Bayesian framework. 
In Section~\ref{sec:datq}, we apply our framework to observed exoplanetary systems.

@@

\begin{figure*}[ht!]
    \script{coordinate.py}
    \gridline{
        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system. The grey circle corresponds to a constant $\psi$ value.}
        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system. The grey circle corresponds to a constant $i_\star$ value.}
    }
    \vspace*{-1.5cm}
    \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.}
    \label{fig:coord}
\end{figure*}

@@

\begin{figure*}[ht!]
    \script{transform.py}
    \includegraphics[width=\linewidth]{figures/transform.pdf}
    \caption{Distribution transformation.}
    \label{fig:transform}
\end{figure*}

>>>>

Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation environment and dynamical evolution of a planetary system. In exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured. To find the true stellar obliquity, information about the stellar inclination $i_\star$ is required.
Here we show that, while it is important to know the stellar inclination to tightly constrain the stellar obliquity of an individual system, the stellar inclination is not necessarily required for the population stellar obliquity inference. Stellar obliquity distribution can be inferred purely from sky-projected stellar obliquities.
We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems from their projected stellar obliquities, stellar inclinations, or both.
Applying the framework to all systems with projected stellar obliquity measurements, which are mostly Hot Jupiter systems, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees. Misaligned systems have nearly isotropic stellar obliquity distribution, at odds with previous observations.

@@

Stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$. 
The angle is an important trace of the formation environment and dynamical evolution of a planetary system. The stellar obliquity evolution can be summarized into three phases. First, the formation and evolution of a protoplanetary disk around its host star determine the primordial stellar obliquity. Second, post-formation dynamical evolution in the planetary system excites the mutual inclinations between planets or stellar companions and further sculpts the stellar obliquity. Lastly, tidal realignment of the host star's spin axis to the planet's orbital axis could reduce the stellar obliquity, if the process is efficient.

In reality, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured. Stellar obliquity $\psi$ can be derived if the stellar inclination $i_\star$ is also known. The calculation follows \citep[e.g.,][]{Fabrycky09}:
\begin{equation}\label{eqn:psi}
    \cos{\psi} = \sin{i_\star}\sin{i_{\rm orb}}\cos{\lambda} + \cos{i_\star}\cos{i_{\rm orb}},
\end{equation}
where $i_{\rm orb}$ is the angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the angle between $\bf{n}_{\star}$ and the observer's line of sight.
The stellar inclination can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology. Since the stellar inclination measurement is degenerate between $i_\star$ and $180\degr - i_\star$, $i_\star$ is usually assumed to be less than $90\degr$.
Since the sky-projected stellar obliquities can be measured only on transiting exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.

Stellar obliquity is important to understand; however, not all systems have a stellar inclination measurement.
For systems without measured $i_\star$, it is still possible to infer the stellar obliquity $\psi$ from the sky-projected obliquity $\lambda$ by assuming the stellar inclination $i_\star$ follows the isotropic distribution, although results in a larger uncertainty on $\psi$ \citep{Fabrycky09}.

In this work, we extend the inference of stellar obliquity to a population. We identify an approach to infer $\psi$ distribution from $\lambda$ distribution and show that a similar constraint can be reached with or without stellar inclination measurement.
In Section~\ref{sec:jacobian}, we present the transformation between the stellar obliquity distribution and the sky-projected stellar obliquity distribution.
In Section~\ref{sec:hbm}, we introduce a hierarchical Bayesian framework to infer the population stellar obliquity distribution. 
And lastly, in Section~\ref{sec:data}, we apply our framework to all exoplanetary systems with measured sky-projected obliquity.

\begin{figure*}[ht!]
    \script{coordinate.py}
    \gridline{
        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system. The grey circle corresponds to a constant $\psi$ value.}
        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system. The grey circle corresponds to a constant $i_\star$ value.}
    }
    \vspace*{-1.5cm}
    \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$. Here we define the observer's line of sight as one of the two horizontal axes, and the orbital axis of the planet as the vertical axis. We approximate the orbital inclination to $90\degr$.}
    \label{fig:coord}
\end{figure*}

@@

@@

\begin{figure*}[ht!]
    \script{transform.py}
    \includegraphics[width=\linewidth]{figures/transform.pdf}
    \caption{Distribution transformations between the stellar obliquity $\cos{\psi}$ (1st column), sky-projected stellar obliquity $\lambda$ (2nd column), and stellar inclination $i_\star$ (3rd column). The inferred $\cos{\psi}$ assuming an isotropic stellar inclination is shown in the 4th column. Each row has a simulated $\cos{\psi}$ distribution. Assuming a uniform $\theta$, the simulated $\lambda$ and $i_\star$ distributions are shown in grey histograms, and the analytical solutions are shown as blue curves.}
    \label{fig:transform}
\end{figure*}
